---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# parn94: Implementation of Richard Parncutt's Psychoacoustic Harmony Algorithms

[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)

`parn94` is an R package that implements Richard Parncutt's 
psychoacoustic harmony algorithms,
as described in Parncutt & Strasburger (1994).

## Installation

You can install the current version of `parn94` from Github
by entering the following commands into R:

``` r
if (!require(devtools)) install.packages("devtools")
devtools::install_github("parn94")
```

## Example usage

This package provides a variety of methods for analysing chords
in isolation and in combination.

Most functions can be used with numeric inputs, 
which will be interpreted as MIDI note numbers.
Methods are also provided for various chord classes in the 
`hrep` package, such as `pi_chord`, `pi_sparse_spectrum`, and so on.
These inputs are internally coerced to `pi_sparse_spectrum`
before continuing the analysis.

Key functions include:

- `pure_sonor()` - returns the *pure sonorousness* of a sonority the amount of pitch content in a sonority,
corresponding to the audibility of its pure tone components
after accounting for auditory masking.
- `complex_sonor()` - returns the *complex sonorousness* of a sonority,
the extent to which a sonority resembles a harmonic series.
- `multiplicity()` - estimates *multiplicity*,
how many tones are perceived in a sonority.
- `pitch_commonality()` - estimates the *pitch commonality* 
of a pair of sonorities.
- `pitch_distance()` - estimates the *pitch distance*
between a pair of sonorities.

```{r}
library(parn94)

c_maj <- c(60, 64, 67) # C major triad
c_dim <- c(60, 63, 66) # C diminished triad
g_maj <- c(59, 62, 67) # G major triad

# Pure sonorousness
pure_sonor(c_maj)
pure_sonor(c_dim)

# Complex sonorousness
complex_sonor(c_maj)
complex_sonor(c_dim)

# Multiplicity
multiplicity(c_maj)
multiplicity(c_dim)

# Pitch commonality
pitch_commonality(c_maj, g_maj)
pitch_commonality(c_maj, c_dim)

# Pitch distance
pitch_distance(c_maj, g_maj)
pitch_distance(c_maj, c_dim)
```

## References

Parncutt, R., & Strasburger, H. (1994). Applying psychoacoustics in composition: “Harmonic” progressions of “nonharmonic” sonorities. Perspectives of New Music, 32(2), 88–129.
